---
title: "Homework 6"
author: "Kevin Chen (914861432) John Mayhew (914807483)"
date: "5/31/2020"
output:
  pdf_document: default
  html_document: default
---
```{r, include = FALSE}
library("quantmod")
library("YieldCurve")
```

## 1. Obtain the monthly data of interest rate (aka. annual yield) of different U.S. treasury bonds with maturities 1-month, 3-month, 6-month, 1-year, 2-year, 3-year, 5-year, 7-year, 10-year, 20-year, and 30-years.

```{r, echo = FALSE}
rm(list=ls())
library("quantmod")
getSymbols(c("GS1M","GS3M","GS6M","GS1","GS2","GS3","GS5","GS7","GS10","GS20","GS30"), src = "FRED")
```

NOT SURE HOW/WHAT TO DISPLAY FOR THIS, MAYBE HEAD OF EACH BOND?

## 2. Find the (annual) yield of a 3-month T-bill in March 2007. Then find the (annual) yield of 20-year T-bond in the same month. (Find those numbers from the time series you loaded into R in the last question. Note that the unit of the yield data stored in R is percentage point. That is if you get a 4 from your data it means a 4% or 0.04 annual yield.) What would you comment on those numbers?

```{r, echo = FALSE}
date = "2007-03-01"
Trates<- cbind(GS3M[index(GS3M)==date],GS20[index(GS20)==date])
Trates
```

There is a slightly higher yield for a 3 month T-bill in March 2007 than a 20 year T-bond. This means the return on the 3 month T-bill is higher than the reteun on the 20 year T-bond.

## 3. Now, draw a scatter plot of the yield data of all bonds in March 2007 against their maturity dates (you will have 11 data points in the scatter plot).

```{r, echo = FALSE}
Trates2 <-cbind(GS1M[index(GS1M)==date],GS3M[index(GS3M)==date],GS6M[index(GS6M)==date],
              GS1[index(GS1)==date],GS2[index(GS2)==date],GS3[index(GS3)==date],
              GS5[index(GS5)==date], GS7[index(GS7)==date],GS10[index(GS10)==date],
              GS20[index(GS20)==date],GS30[index(GS30)==date])
maturity<-c(1/12,3/12,6/12,1,2,3,5,7,10,20,30)
plot(maturity,Trates2, main=paste("Yield Curve in ", date, sep=""))
```

## 4. Fit a yield curve using the Nelson-Siegel model for March 2007. Interpret your $\beta_0$ and $\beta_1$ estimates. Also, find the sign of your $\beta_2$ estimate. What does the sign tell you? 

```{r, echo = FALSE}
NSresults <- Nelson.Siegel(rate = Trates2, maturity = maturity)
NSresults
beta0 <- NSresults[,1]
beta1 <- NSresults[,2]
beta2 <- NSresults[,3]
```

$\beta_0$ is the long-term interest rate. $\beta_1$ is the long-to-short-term spread, which means the  difference between the long term interest rate and the short term interest rate. $\beta_2$ is the curvature parameter and since $\beta_2$ is less than 0, the curve produces a trough, which means it is inverted.

## 5. What does your Nelson-Siegel model predict about the annual yield of U.S. treasury bonds in the secondary market in March 2007 that has 2.5 years left to maturity?

```{r, echo = FALSE}
predict <- NSrates(NSresults, 2.5)
predict
```

Our Nelson-Siegel model predicts `r predict`% as the annual yield of a US treasury bond in the secondary market in March 2007 with 2.5 years left to maturity. 

## 6. 6. Suppose there is a zero-coupon bond with face value $1000 that in March 2007 has 2.5 years left to maturity. Calculate the predicted price of this zero-coupon bond using your predicted yield in the last question.

```{r, echo = FALSE}
fv = 1000
price <- fv/((1+0.04573102)^2.5)
price
```

### Appendix 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```